---
title: "Assignment 6: GLMs (Linear Regressios, ANOVA, & t-tests)"
author: "Yikai Jing"
output: pdf_document
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: console
---

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on generalized linear models. 

## Directions
1. Change "Student Name" on line 3 (above) with your name.
2. Work through the steps, **creating code and output** that fulfill each instruction.
3. Be sure to **answer the questions** in this assignment document.
4. When you have completed the assignment, **Knit** the text and code into a single PDF file.
5. After Knitting, submit the completed exercise (PDF file) to the dropbox in Sakai. Add your last name into the file name (e.g., "Fay_A06_GLMs.Rmd") prior to submission.

The completed exercise is due on Monday, February 28 at 7:00 pm.

## Set up your session 
1. Set up your session. Check your working directory. Load the tidyverse, agricolae and other needed packages. Import the *raw* NTL-LTER raw data file for chemistry/physics (`NTL-LTER_Lake_ChemistryPhysics_Raw.csv`). Set date columns to date objects.

2. Build a ggplot theme and set it as your default theme.

```{r setup}
#1
getwd()
library(tidyverse)
library(agricolae)
library(lubridate)
water <- read.csv('~/Environmental_Data_Analytics_2022/Data/Raw/NTL-LTER_Lake_ChemistryPhysics_Raw.csv', stringsAsFactors = TRUE)

water$sampledate <- as.Date(water$sampledate, "%m/%d/%y")
class(water$sampledate)

#2
mytheme <- theme_classic(base_size = 12) +
  theme(axis.text = element_text(color = "black"),
        legend.position = "top")

```

## Simple regression
Our first research question is: Does mean lake temperature recorded during July change with depth across all lakes?

3. State the null and alternative hypotheses for this question:
> Answer:
H0: mean lake temperature recorded during July DOES NOT change with depth across all lakes.
Ha: mean lake temperature recorded during July DOES change with depth across all lakes.


4.  Wrangle your NTL-LTER dataset with a pipe function so that the records meet the following criteria: 
 * Only dates in July. 
 * Only the columns: `lakename`, `year4`, `daynum`, `depth`, `temperature_C`
 * Only complete cases (i.e., remove NAs)

5. Visualize the relationship among the two continuous variables with a scatter plot of temperature by depth. Add a smoothed line showing the linear model, and limit temperature values from 0 to 35 °C. Make this plot look pretty and easy to read.

```{r scatterplot}
#4
water.processed <-
  water %>%
  mutate(Month = month(sampledate)) %>%
  filter(Month == "7") %>%
  select(lakename, year4, daynum, depth, temperature_C) %>%
  drop_na(depth) %>%
  drop_na(temperature_C)

#5
graph1 <-
  ggplot(water.processed, aes(x = depth, y = temperature_C)) +
  geom_point() +
  geom_smooth(method = lm) +
  ylim(0,35)+
  ggtitle("Graph 1: Scatterplot of Temperature vs depth") +
  ylab("Temperature (Celsius)") +
  xlab("Depth (meter)") +
  mytheme
print(graph1)

```


6. Interpret the figure. What does it suggest with regards to the response of temperature to depth? Do the distribution of points suggest about anything about the linearity of this trend?

> Answer: The detected temperature of the lake tend to be higher with a sjallow depth and tend to decrease as depth of the detection point increases. However, we cannot confirm that this is a linear relationship because the distribution of the scatter points are not very linear.


7. Perform a linear regression to test the relationship and display the results

```{r linear.regression}
#7
temp.regression <- lm(data =water.processed, temperature_C ~ depth)
summary(temp.regression)

```


8. Interpret your model results in words. Include how much of the variability in temperature is explained by changes in depth, the degrees of freedom on which this finding is based, and the statistical significance of the result. Also mention how much temperature is predicted to change for every 1m change in depth. 

> Answer:With an R-squared value of 0.7387, about 73.87% of the observed variance is explained by this model. With a degree of freedom of 9726 and p value less than 2.2e-16, we can successfully reject the null hypothesis and state that mean lake temperature recorded during July DOES change with depth across all lakes. For every 1m change in the depth, temperature would change for -1.94621 degree Celsius (for a negative relationship).


---

## Multiple regression
Let's tackle a similar question from a different approach. Here, we want to explore what might the best set of predictors for lake temperature in July across the monitoring period at the North Temperate Lakes LTER. 


9. Run an AIC to determine what set of explanatory variables (year4, daynum, depth) is best suited to predict temperature.

10. Run a multiple regression on the recommended set of variables. 

```{r temperature.model}
#9
temp.AIC <- lm(data =water.processed, temperature_C ~ depth + year4 + daynum)
step(temp.AIC)

#10
TPmodel <- lm(data = water.processed, temperature_C ~ depth + year4 + daynum)
summary(TPmodel)

```

11. What is the final set of explanatory variables that the AIC method suggests we use to predict temperature in our multiple regression? How much of the observed variance does this model explain? Is this an improvement over the model using only depth as the explanatory variable?

> Answer: The AIC resulTs suggest that none of the three explanatory variables would generate a better combination for the multi-linear regression. The final set of explanatory variables suggested by the AIC include depth, year4, and daynum.This model explains 74.12% of the observed variance, which is slightly (0.25%) higher than the model using only depth.



---
## Analysis of Variance

12. Now we want to see whether the different lakes have, on average, different temperatures in the month of July. Run an ANOVA test to complete this analysis. (No need to test assumptions of normality or similar variances.) Create two sets of models: one expressed as an ANOVA models and another expressed as a linear model (as done in our lessons).

```{r anova.model}
#12
# Format as aov
water.anova1 <- aov(data = water.processed, temperature_C ~ lakename)
summary(water.anova1)

# Format as lm
water.anova2 <- lm(data = water.processed, temperature_C ~ lakename)
summary(water.anova2)
```

13. Is there a significant difference in mean temperature among the lakes? Report your findings. 

> Answer: Yes. Based on the summary results of both of water.anova1 model, we can conclude that there is a significant difference in mean temperature among the lakes with a significance value less than 0.001. The water.anova2 model demonstrated similar findings with almost all lakes having different means at significance level of 0 except for Crampton Lake having a difference in mean temperature at significance level of 0.001. We received a R-squared value of 0.03953, meaning that approximately 4% of variation in the response is explained by the water.anova2 model.



14. Create a graph that depicts temperature by depth, with a separate color for each lake. Add a geom_smooth (method = "lm", se = FALSE) for each lake. Make your points 50 % transparent. Adjust your y axis limits to go from 0 to 35 degrees. Clean up your graph to make it pretty. 

```{r scatterplot.2}
#14.
graph2 <-
  ggplot(water.processed, aes(x = depth, y = temperature_C, color = lakename)) +
  geom_point(alpha =0.5) +
  geom_smooth(method = lm, se = FALSE) +
  ylim(0,35)+
  ggtitle("Graph 2: Scatterplot of Temperature vs depth by lake") +
  xlab("Depth (meter)") +
  ylab("Temperature (Celsius)") +
  mytheme
print(graph2)

```

15. Use the Tukey's HSD test to determine which lakes have different means.

```{r tukey.test}
#15
TukeyHSD(water.anova1)
water.groups <- HSD.test(water.anova1, "lakename", group = TRUE)
water.groups
```

16.From the findings above, which lakes have the same mean temperature, statistically speaking, as Peter Lake? Does any lake have a mean temperature that is statistically distinct from all the other lakes?

>Answer:Ward Lake, Peter Lake and Paul Lake have statistically same mean temperature. No lake has a mean temperature that is statistically distinct from all the other lakes.


17. If we were just looking at Peter Lake and Paul Lake. What's another test we might explore to see whether they have distinct mean temperatures? 

>Answer: We may conduct a Two Sample t test for Comparing Two Means of the Peter Lake and Paul lake solely.


